\section{Cloud File Synchronisation Services}\label{sec:application:cloud_file_synchronisation}

\newcommand{\dropbox}{\emph{DropBox}\xspace}
\newcommand{\imageFileSize}{\ensuremath{S_I}\xspace}
\newcommand{\imageArrivalRate}{\ensuremath{\lambda_I}\xspace}
\newcommand{\panTransferRate}{\ensuremath{B_{P}}\xspace}
\newcommand{\interarrivaltime}{\ensuremath{t_I}\xspace}
\newcommand{\clientpreparationtime}{\ensuremath{C}\xspace}
\newcommand{\uploadbandwidth}{\ensuremath{B_U}\xspace}
\newcommand{\idleThreshold}{\ensuremath{\tau}\xspace}
\newcommand{\serverpreparationtime}{\ensuremath{S}\xspace}
\newcommand{\downloadbandwidth}{\ensuremath{B_D}\xspace}
\newcommand{\startupDelay}{\ensuremath{\sigma}\xspace}
\newcommand{\sojournTime}{\ensuremath{\Sigma}\xspace}
\newcommand{\relativeDisconnectedTime}{\ensuremath{\Delta}\xspace}
\newcommand{\connectionCount}{\ensuremath{K}\xspace}
\newcommand{\algointerval}{\emph{Interval}\xspace}
\newcommand{\algosize}{\emph{Size}\xspace}
\newcommand{\algoimmediate}{\emph{Immediate}\xspace}
\newcommand{\numberOfFiles}{\ensuremath{n}}

%file storage
While the previous sections studied video playback, this is far from the only application deployed in today's networks.
Another prominent type are file synchronisation services.
They are offered by a multitude of providers, e.g. Google Drive, and Microsoft OneDrive, and the current leader in the field Dropbox.
A general study of the \gls{QoE} influence factors of file storage services is undertaken in~\cite{Qian2011b}.
Here, the authors provide a model allowing for the evaluation of cloud service providers according to a variety of metrics, including bandwidth, latency, and response time.
Due to its popularity, this section focuses on Dropbox.

In~\cite{Amrehm2013} the authors study the main impact factors on \gls{QoE} of Dropbox users.
They find that the main impact factor for user \gls{QoE} is the waiting time for file synchronisation.
Thus, this is the main factor the cloud storage operator will dimension the service for, in order to ensure a high user satisfaction. 
However, the operator of the file synchronisation service is not the only stakeholder in this scenario.
As in earlier sections, the signalling traffic imposed by application traffic is of concern to network providers.
Furthermore, the services users are, as discussed in \refchap{chap:network}, interested in reducing the power drain due to network connectivity in order to increase their devices battery life.

In \refsec{sec:application:cloud_file_synchronisation:system_model} we first propose a model in order to evaluate the Dropbox service, define relevant metrics for each stakeholder as well as different scheduling algorithms used to trigger file synchronisation.
Then, in \refsec{sec:application:cloud_file_synchronisation:application_measurements} we perform network measurements using the PlanetLab~\cite{Chun2003} research network to be used as input for our model.
Finally, in \refsec{sec:application:cloud_file_synchronisation:numerical_evaluation} we evaluate our model using a non-stationary discrete event simulation for all considered metrics and scheduling algorithms.

\input{application/cloud_file_synchronization/system_model/system_model}
\input{application/cloud_file_synchronization/application_measurements/application_measurements}
\input{application/cloud_file_synchronization/numerical_evaluation/numerical_evaluation}

